### 2021-01-18 BMC-Journal of Biomedical Semantics/submission 0/bmc_template/methodology.tex
method 5

In order to process the EMR schema set $S$, we use the methodology shown in \textbf{figAlgorithm}. For each data representation $s_i$, we first serialize its tables into Comma Separated Value (CSV) formatted flat files. Each table has a separate file, which contains the attribute names in the header and the data instances (following the same format and CSV struture), in subsequent lines. Each file is loaded in memory and converted into an algorithmic iterable structure. For each attribute, we then generate the suffix tree, which provides all possible substring representations contained within the attribute name. Then for each node of this suffix tree, if its length is greater than 2 (to avoid lookups for alphabets), we then query UMLS to retrieve its associated semantic concepts. \textbf{threshold}
If atleast one semantic concept is found for the substring, we replace the substring with the \textbf{most relevant} concept, creating the Suffix Concept Tree (SCT). The SCT is structurally smaller than or equal to the suffix tree, since all suffix which are smaller than 2 and those without a qualifying semantic concept, are discarded. Finally for each node of the SCT, the substring from the suffix tree and its corresponding concept is placed into a sentence, which is converted into an embedding vector using \textbf{BioBERT}. This embedding vector represents the domain qualified semantics of the source attribute.

@@

Meanwhile the process continues for the next attribute, then the next table, and finally the next system, till no further processing is possible. The flat enriched schema, following the design shown in \textbf{figenrichedSchemaPic} provides the input to our knowledge base. The logical structure of the enriched attribute, is represented in \textbf{figAttributeRepresentation}. Here, the elements of this particular data-structure are categorized into three parts. ``AttributeContext" contains the metadata conforming to an instance of the attribute's existence,  with its name, container table name and schema name, acting as a pointer, and schema version, source, and recorded date providing the version control features. In this manner, the attribute's fully qualified reference, along with its existential context is identified. ``AttributeType" then encapsulates the data type features of this attribute's instances, with its primitive data type, and values providing a representation of the attribute's instance. Finally, ``Attribute Semantics" holds, the semantic information of this attribute, in the form of its suffix tree, SCT and Tree Embedding.
A pair of this data-structure with distinct references, are then used as an input to the schema map generation process, which is explained in the next step.

>>>>

In order to process the EMR schema set $S$, we use the methodology shown in \textbf{figAlgorithm}. For each data representation $s_i$, we first serialize its tables into Comma Separated Value (CSV) formatted flat files. Each table has a separate file, which contains the attribute names in the header and the data instances (following the same format and CSV struture), in subsequent lines. Each file is loaded in memory and converted into an algorithmic iterable structure. For each attribute, we then generate the suffix tree, which provides all possible substring representations contained within the attribute name. Then for each node of this suffix tree, if its length is greater than 2 (to avoid lookups for alphabets), we then query UMLS with approximate search strategy.Using Lexical Variant Generation (LVG) this strategy retrieves an expanded list of associated semantic concepts.
If atleast one semantic concept is found for the substring, we replace the substring with the most relevant concept, based on the highest value of the confidence measure, creating the Suffix Concept Tree (SCT). The SCT is structurally smaller than or equal to the suffix tree, since all suffix which are smaller than 2 and those without a qualifying semantic concept, are discarded. Finally for each node of the SCT, the substring from the suffix tree and its corresponding concept is placed into a sentence, which is converted into an embedding vector using \textbf{BioBERT}. This embedding vector represents the domain qualified semantics of the source attribute.

@@

Meanwhile the process continues for the next attribute, then the next table, and finally the next system, till no further processing is possible. The flat enriched schema, following the design shown in \textbf{figEnrichedSchema} provides the input to our knowledge base. The logical structure of the enriched attribute, is represented in \textbf{figAttributeRepresentation}. Here, the elements of this particular data-structure are categorized into three parts. ``AttributeContext" contains the metadata conforming to an instance of the attribute's existence,  with its name, container table name and schema name, acting as a pointer, and schema version, source, and recorded date providing the version control features. In this manner, the attribute's fully qualified reference, along with its existential context is identified. ``AttributeType" then encapsulates the data type features of this attribute's instances, with its primitive data type, and values providing a representation of the attribute's instance. Finally, ``Attribute Semantics" holds, the semantic information of this attribute, in the form of its suffix tree, SCT and Tree Embedding.
A pair of this data-structure with distinct references, are then used as an input to the schema map generation process, which is explained in the next step.
